---
title: "MNIST"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### MNIST dataset - Logistic regression
```{r include=FALSE}
library(dslabs)
library(ggplot2)
library(caret)
library(dplyr)
```

Classification of mnist data for digits 2 and 7

Load the data
```{r}
data("mnist_27")
mnist_27$train %>% ggplot(aes(x_1, x_2, color = y)) + geom_point()
```

Create logistic regression model
```{r}
fit <- glm(y ~ x_1 + x_2, data=mnist_27$train, family="binomial")
```

Predict with the glm model and classification
```{r}
p_hat <- predict(fit, newdata = mnist_27$test)
y_hat <- factor(ifelse(p_hat > 0.5, 7, 2)) 
```

Confusion matrix
```{r}
confusionMatrix(data = y_hat, reference = mnist_27$test$y) 
```

Plot true probabilities
```{r}
mnist_27$true_p %>% ggplot(aes(x_1, x_2, z=p, fill=p)) +  geom_raster() + 
  scale_fill_gradientn(colors=c("#F8766D","white","#00BFC4")) +  
  stat_contour(breaks=c(0.5),color="black")
```

Predict the probabilities with Logistic Regression and Plot predicted probabilities
```{r}
p_hat <- predict(fit, newdata = mnist_27$true_p) 

mnist_27$true_p %>% mutate(p_hat = p_hat) %>% 
  ggplot(aes(x_1, x_2,  z=p_hat, fill=p_hat)) + geom_raster() +   
  scale_fill_gradientn(colors=c("#F8766D","white","#00BFC4")) +   
  stat_contour(breaks=c(0.5),color="black") 
```

Plot Logistic regression line with test data
```{r}
mnist_27$true_p %>% mutate(p_hat = p_hat) %>% ggplot() +   
  stat_contour(aes(x_1, x_2, z=p_hat), breaks=c(0.5), color="black") +    
  geom_point(mapping = aes(x_1, x_2, color=y), data = mnist_27$test) 
```

Logistic regression is able to classify the data with 76% accuracy.

